---
title: "US Conference Board Consumer Confidence"
author: "pgg1309"
date: "2022-02-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE,
    include = TRUE
)
require(fpp3)
require(tidyverse)
```

```{r functions}
plot_var <- function(db, id, sdate = "2007 Jan") {
  
  x <- filter(db, series_id %in% id) %>%
    group_by_key() %>% 
    mutate(chg = value - lag(value)) %>% 
    drop_na() %>% 
    group_by_key() %>% 
    mutate(trend = signal::sgolayfilt(value, p = 1, n = 7))
  
  # preservou estrutura de grupos
  xlast <- x %>% slice_tail()

  x %>% 
    filter_index(sdate ~ .) %>% 
  ggplot(aes(x=month)) +
  geom_line(aes(y = value), col = 'darkgrey', size = 1) + 
  geom_line(aes(y = trend), size = 1, col = 'blue') +
  geom_point(aes(y = value), data = xlast, size = 2) +  
  scale_y_continuous(
    breaks = scales::breaks_pretty(n = 8),
    sec.axis = dup_axis(name = NULL)
  ) +
  scale_x_yearmonth(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 12)
  ) +
  labs(
    x = NULL,
    y = "index"
  ) +
  theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom") +
  facet_wrap(vars(series_name), dir = 'v', scales = 'free_y')
  
}
```


## Main takeaways
*Obs: the cutoff date for the preliminary result was Feb 16*.

- Consumer confidence slightly down in February, with diverging components: present situation slightly up and expectations slightly down.
- Assessment of the labor market worsened marginally but overall it still shows a very strong labor market.


```{r}
# read data
db <- readRDS("E:/RProjs/us-monthly-db/data/conferenceboard-data.rds")


```

## Consumer confidence
```{r}
plot_var(db, "USCNFCONQ")
```

## Present situation
```{r}
plot_var(db, "USCNPSIT")
```

## Expectations index
```{r}
plot_var(db, "USCNEXPT")
```

## Labor market
```{r}
plot_var(db, c("USTCBEPN", "USTCBEPP"))
```

```{r}
# read unemployment rate
myboard <- pins::board_folder("E:/RData/CPS")
ur <- pins::pin_read(myboard, 'data.1.AllData')
ur <-
  filter(ur, stringr::str_trim(`series_id        `) == "LNS14000000") %>%
  transmute(month = yearmonth(paste0(year, period)),
            series_id = "LNS14000000",
            value = parse_number(`       value`),
            series_name = "US Unemployment Rate") %>% 
  as_tsibble(index = month, key = series_id)


labor_db <- filter(db, series_id %in% c("USTCBEPN", "USTCBEPP")) %>% 
  select(-series_name) %>% 
  as_tibble() %>% 
  pivot_wider(names_from = series_id, values_from = value) %>% 
  mutate(`Jobs hard - plentiful` = USTCBEPP - USTCBEPN) %>% 
  left_join(select(ur, month, value), by = 'month') %>% 
  rename(`Unemployment rate` = value) %>% 
  as_tsibble() 

labor_model <- labor_db %>% 
  model(
    arima = ARIMA(`Unemployment rate` ~ PDQ(0,0,0) + `Jobs hard - plentiful`)
  )

extremos <- augment(labor_model) %>%
  drop_na() %>% 
  filter(
    .innov < quantile(.innov, 0.25) - 3 * IQR(.innov) |
    .innov > quantile(.innov, 0.75) + 3 * IQR(.innov)
  )

# add dummies
labor_db_miss <- labor_db %>% 
  anti_join(extremos) %>% 
  fill_gaps() 


labor_model_miss <- labor_db_miss %>% 
  model(
    arima = ARIMA(`Unemployment rate` ~ PDQ(0,0,0) + `Jobs hard - plentiful`)
  ) 
  
#report(labor_model_miss)

#labor_model_miss %>% gg_tsresiduals()


forecast(labor_model_miss, labor_db) %>% 
  left_join(select(ur, month, value)) %>% 
  ggplot(aes(x = month)) +
  geom_line(aes(y = value, color = 'Unemployment rate'), size = 1) + 
  geom_line(aes(y = .mean, color = 'Fitted'), size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 15)
  ) +
  labs(
    title = "US Conf. Board: Unemployment vs Job hard minus plentiful",
    subtitle = "UR = f(Jobs Hard to Find - Jobs Plentiful)",
    color = NULL,
    x = NULL,
    y = "%"
  ) +
  theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom")
  
  



```

## Comparing consumer confidence measures
```{r}
# read michigan
require(DatastreamDSWS2R)
mydsws <- dsws$new()
michigan <- mydsws$timeSeriesRequest(instrument = "USUMCONSH",
                       datatype = "X",
                       startDate = "1980-01-01",
                       endDate = Sys.Date(),
                       frequency = "M")

db_consumer_measures <- tibble(month = zoo::index(michigan), value = as.vector(michigan)) %>% 
  transmute(month = yearmonth(month),
            series_id = "USUMCONSH",
            value,
            series_name = "US Univ of Michigan Consumer Sentiment") %>% 
  as_tsibble(index = month, key = series_id) %>% 
  bind_rows(filter(db, series_id == "USCNFCONQ"))

plot_var(db_consumer_measures, c("USUMCONSH", "USCNFCONQ"))

```

