---
title: "US Empire State Manufacturing"
author: "Dr. Paulo Gustavo Grahl, CFA"
date: "2022-02-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE,
    include = TRUE
)
require(fpp3)
```

```{r}
# Functions
plot_var <- function(db, id) {
  
  x <- filter(db, series_id %in% c(id, stringr::str_replace(id, "CDISA$", "FDISA"))) %>%
    group_by_key() %>% 
    mutate(chg = value - lag(value)) %>% 
    drop_na() %>% 
    group_by_key() %>% 
    mutate(trend = signal::sgolayfilt(value, p = 1, n = 7))
  
  # preservou estrutura de grupos
  xlast <- x %>% slice_tail()

  x %>% 
    filter_index("2010 Jan" ~ .) %>% 
  ggplot(aes(x=month)) +
  geom_line(aes(y = value), col = 'darkgrey', size = 1) + 
  geom_line(aes(y = trend), size = 1, col = 'blue') +
  geom_point(aes(y = value), data = xlast, size = 2) +  
  scale_y_continuous(
    breaks = scales::breaks_pretty(n = 8),
    sec.axis = dup_axis(name = NULL)
  ) +
  scale_x_yearmonth(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 12)
  ) +
  labs(
    x = NULL,
    y = "diffusion index"
  ) +
  theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom") +
  facet_wrap(vars(series_name), dir = 'v', scales = 'free_y')
  
}
```


## Main takeaways

* Business activity little changed in February, up four points to 3.1 but below
market expectations of 12.
* New orders and shipments marginally up, but at a level that implies a material 
slowdown in activity compared to 4Q 2021.
* Unfilled orders, delivery time, and inventories continue to show the impact 
of supply chain bottlenecks.
* Prices paid remained at high level and prices received reached a new high -- not
good for inflation.
* Employment remains strong.Also capex and tech spending plans remain close to the highs.


## Charts
```{r}
# Read data
empireboard <- pins::board_folder("E:/RData/EMPIREMANUF")
empire <- pins::pin_read(empireboard, 'empire-manufacturing')
```

******
### General Business Conditions
```{r}
plot_var(empire, 'GACDISA')
```

******
### Empire Manufacturing - modified
```{r}
empiremod <- bind_rows(
  filter(empire, series_id %in% c('IVCDISA', 'NOCDISA', 'NECDISA', 'SHCDISA', 'DTCDISA')) %>% 
  index_by() %>% 
  summarize(value = mean(value)/2+50, series_name = 'Current Modified') %>% 
  update_tsibble(key = series_name)
,
filter(empire, series_id %in% c('IVFDISA', 'NOFDISA', 'NEFDISA', 'SHFDISA', 'DTFDISA')) %>% 
  index_by() %>% 
  summarize(value = mean(value)/2+50, series_name = 'Future Modified') %>%
  update_tsibble(key = series_name)
) %>% 
   group_by_key() %>% 
    mutate(trend = signal::sgolayfilt(value, p = 1, n = 7))

empiremodlast <- slice_tail(empiremod)

 empiremod %>% 
   filter_index("2010 Jan" ~ .) %>% 
  ggplot(aes(x=month)) +
  geom_line(aes(y = value), col = 'darkgrey', size = 1) + 
  geom_line(aes(y = trend), size = 1, col = 'blue') +
  geom_point(aes(y = value), data = empiremodlast, size = 2) +  
  scale_y_continuous(
    breaks = scales::breaks_pretty(n = 8),
    sec.axis = dup_axis(name = NULL)
  ) +
  scale_x_yearmonth(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 12)
  ) +
  labs(
    x = NULL,
    y = "diffusion index"
  ) +
  theme_bw() +
    theme(panel.border = element_blank(), legend.position = "bottom") +
  facet_wrap(vars(series_name), dir = 'v', scales = 'free_y')

```

******
### New Orders
```{r}
plot_var(empire, 'NOCDISA')
```

******
### Shipments
```{r}
plot_var(empire, 'SHCDISA')
```

******
### Unfilled Orders
```{r}
plot_var(empire, 'UOCDISA')
```

******
### Delivery Time
```{r}
plot_var(empire, 'DTCDISA')
```

******
### Inventories
```{r}
plot_var(empire, 'IVCDISA')
```

******
### Prices Paid
```{r}
plot_var(empire, 'PPCDISA')
```

******
### Prices Received
```{r}
plot_var(empire, 'PRCDISA')
```

******
### Number of Employees
```{r}
plot_var(empire, 'NECDISA')
```

******
### Average Employee Workweek
```{r}
plot_var(empire, 'AWCDISA')
```

******
### Capital Expenditures
```{r}
plot_var(empire, 'CECDISA')
```

******
### Technology Spending
```{r}
plot_var(empire, 'TSCDISA')
```

