---
title: "sandp-earnings"
author: "pgg1309"
date: "2022-05-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9
)
```



# Introduction

```{r}
library(fpp3)
library(DatastreamDSWS2R)

### Global variables
update.data <- TRUE

### Functions
get_data <-
  function(instrument,
           type,
           sdate = "1985-01-01",
           edate = Sys.Date(),
           freq = "D",
           update = FALSE) {
    if (update) {
      mydsws <- dsws$new()
      mydsws$chunkLimit <- 50L
      xx <- mydsws$timeSeriesRequest(
        instrument = instrument,
        datatype = type,
        startDate = sdate,
        endDate = edate,
        frequency = freq
      )
      xx <- tibble(index = as.Date(zoo::index(xx)),
                        value = as.vector(xx),
      )
      fname  <- paste0(stringr::str_extract_all(instrument, "[:alnum:]")[[1]], collapse = "")
      fname2 <- paste0(stringr::str_extract_all(type, "[:alnum:]")[[1]], collapse = "") 
      saveRDS(xx, paste0("data/sandp-earnings/", fname, "-", fname2, ".rds"))
      return(xx)
    } else {
      fname  <- paste0(stringr::str_extract_all(instrument, "[:alnum:]")[[1]], collapse = "")
      fname2 <- paste0(stringr::str_extract_all(type, "[:alnum:]")[[1]], collapse = "") 
      xx <- readRDS(paste0("data/sandp-earnings/", fname, "-", fname2, ".rds"))
      return(xx)
    }
  }


daily2month <- function(db) {
  db %>% 
    mutate(m = yearmonth(index)) %>%
    group_by(m) %>%
    summarize(value = mean(value, na.rm = TRUE),
              nobs = n(), .groups = "drop") %>%
    select(index = m, value, nobs)
  
}

### Loads data


## Daily
eps_ttm <- get_data("@:USSP500", "A12TE", update = update.data)
spx_div <- get_data("S&PCOMP", "IDV", update = update.data)
spx_price <- get_data("@:USSP500", "APRICE", update = update.data)
eps_ftm <- get_data("@:USSP500", "A12FE", update = update.data)
sales_ttm <- get_data("@:USSP500", "AS12TS", update = update.data)
sales_ftm <- get_data("@:USSP500", "AS12FS", update = update.data)
us10y_real <- get_data("TRUSI10", "RY", update = update.data) %>% drop_na()

## Last 12 month avg, Quarterly
gdp_nominal <- get_data('USGDP...B', 'X', freq = 'Q', update = update.data)
gdp_nominal <- gdp_nominal %>% 
  mutate(index = yearquarter(index),
         value = slider::slide_dbl(value, mean, .before = 3, .complete = TRUE)) 
  
gross_output <- get_data('USGOALLIB', 'X', freq = 'Q', update = update.data)
gross_output <- gross_output %>% 
  mutate(index = yearquarter(index),
         value = slider::slide_dbl(value, mean, .before = 3, .complete = TRUE)) 

## Monthly (last 12m avg, monthly)
cpi <- get_data("USCONPRCE", "X", freq = "M", update = update.data)
cpi <- cpi %>% 
  mutate(index = yearmonth(index),
         value = slider::slide_dbl(value, mean, .before = 11, .complete = TRUE))

# convert quarterly to monthly
xx <- tempdisagg::td(
  ts(gdp_nominal$value, start = c(1985, 1), freq = 4)  ~ 1,
  conversion = 'mean',
  to = "monthly", method =  "chow-lin-maxlog")

gdp_monthly <- tsbox::ts_tbl(predict(xx)) %>% 
  transmute(index = yearmonth(time), value)


# select Thu (weekly data)
spx <- bind_rows(
  list(eps = eps_ttm,
       div = spx_div,
       price = spx_price,
       epsf = eps_ftm,
       sps = sales_ttm,
       spsf = sales_ftm),
  .id = 'key') %>% 
  pivot_wider(names_from = key, values_from = value) %>% 
  arrange(index) %>% 
  filter(wday(index) == 5)


```

## Total earnings

The chart below plots total S&P earnings since 1985 in nominal USD. 

```{r fig.width=9, message=FALSE, warning=FALSE}
spx %>% 
  transmute(index, earnings = eps * div) %>% 
  drop_na(earnings) %>% 
  as_tsibble() %>% 
  model(
    lm = TSLM(log(earnings) ~ trend())
  ) %>% 
  augment() %>% 
  ggplot(aes(x=index, y=earnings)) + 
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-6, accuracy = 1),
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +
  
  labs(
    title = "S&P Earnings",
    subtitle = "Trailing 12-months, 2std dev band",
    x = NULL,
    y = 'billion'
  )
```

## Total 'real' earnings
The chart below plots total S&P earnings since 1985 deflated by US CPI
```{r fig.width=9, message=FALSE, warning=FALSE}
real.eps <- spx %>% 
  transmute(index, earnings = eps * div) %>% 
  drop_na(earnings) %>% 
  # select last observation in each month
  group_by(index =yearmonth(index)) %>% 
  slice_tail(n=1) %>% 
  ungroup() %>% 
  as_tsibble() %>% 
  left_join(cpi) %>% 
  drop_na() %>% 
  mutate(real_earnings = earnings / value * value[n()]) %>% 
  model(
    lm = TSLM(log(real_earnings) ~ trend())
  ) %>% 
  augment() 

real.eps %>% 
  ggplot(aes(x=index, y=real_earnings)) + 
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-6, accuracy = 1),
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +
  
  labs(
    title = "Real S&P Earnings",
    subtitle = "Trailing 12-months, 2std dev band",
    x = NULL,
    y = 'billion (constant prices)'
  )
```

## Real earnings growth and 'gap'
O gráfico acima mostra que o lucro do S&P em termos reais cresceu em média no período `r real.eps %>% slice_tail(n=2) %>% mutate(growth = ((.fitted / lag(.fitted))^12 -1)) %>% drop_na(growth) %>%  pull(growth) %>% scales::label_percent(accuracy = 0.1)(.)` ao ano.

O gráfico abaixo mostra o desvio do EPS dessa tendência ('gap') de forma que é possível vislumbrar qual o 'drawdown' do lucro real ao longo dos ciclos. 

```{r}
real.eps %>% 
  mutate(gap = log(real_earnings/.fitted)) %>% 
  ggplot(aes(x=index, y=gap)) + 
  geom_bar(stat = 'identity', alpha = 0.5) +
  geom_line() +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 12),
    sec.axis = dup_axis(name = NULL)
  ) +
  labs(
    title = "Real S&P Earnings",
    subtitle = "Gap from trend",
    x = NULL,
    y = NULL
  )
```


## Earnings vs GDP
```{r fig.width=9}
bind_rows(eps = eps_ttm, div = spx_div, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  transmute(index, value = eps * div) %>% 
  drop_na() %>%  
  daily2month() %>% 
  left_join(gdp_monthly, by = 'index', suffix = c('.earnings', '.gdp')) %>% 
  mutate(earnings_share = value.earnings * 10^(-6) / value.gdp) %>% 
  drop_na(earnings_share) %>% 
  as_tsibble() %>% 
  model(
    lm = TSLM(earnings_share ~ trend())
  ) %>% 
  augment() %>% 
  ggplot(aes(x = as_date(index), y = earnings_share)) +
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = .fitted + 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = .fitted - 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n = 15)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.005),
    labels = scales::label_percent(accuracy = 0.1),
    sec.axis = dup_axis()
  ) +
  labs(
    title = "S&P 500 Earnings",
    subtitle = 'as a share of GDP',
    x = NULL,
    y = NULL
  ) +
  coord_cartesian(ylim = c(0, NA))

```




## Total sales

```{r fig.width=9}
spx %>%
  transmute(index, sales = sps * div) %>% 
  drop_na(sales) %>% 
  as_tsibble() %>% 
  model(lm = TSLM(log(sales) ~trend())) %>% 
  augment() %>% 
  ggplot(aes(x=index, y=sales)) + 
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-6, accuracy = 1),
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +
  
  labs(
    title = "S&P Sales",
    subtitle = "Trailing 12-months, std dev band",
    x = NULL,
    y = 'billion'
  )
```


## Sales vs GDP
```{r fig.width=9}
bind_rows(sps = sales_ttm, div = spx_div, .id = 'key') %>% 
  pivot_wider(names_from = 'key', values_from = value) %>%
  transmute(index, value = sps * div) %>% 
  drop_na() %>% 
  daily2month() %>% 
  left_join(gdp_monthly, by = 'index', suffix = c('.sales', '.gdp')) %>% 
  mutate(sales_share = value.sales * 10^(-6) / value.gdp) %>% 
  drop_na(sales_share) %>% 
  as_tsibble() %>% 
  model(lm = TSLM(sales_share ~trend())) %>% 
  augment() %>% 
  ggplot(aes(x = as_date(index), y = sales_share)) +
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = .fitted + 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = .fitted - 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n = 15)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.025),
    labels = scales::label_percent(accuracy = 1)
  ) +
  labs(
    title = "S&P 500 Sales",
    subtitle = 'as a share of GDP, 2std dev band',
    x = NULL,
    y = NULL
  ) +
  coord_cartesian(ylim = c(0.35, NA))

```



## Margin

```{r fig.width=9}
bind_rows(sps = sales_ttm, eps = eps_ttm, .id = 'key') %>% 
  pivot_wider(names_from = 'key', values_from = value) %>%
  transmute(index, value = eps / sps) %>% 
  drop_na() %>% 
  daily2month() %>% 
  as_tsibble() %>% 
  model(lm = TSLM(value ~ trend())) %>% 
  augment() %>% 
  ggplot(aes(x = as_date(index), y = value)) + 
  geom_line(size = 1) + 
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = .fitted + 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = .fitted - 2*sd(.innov)), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = scales::breaks_extended(n = 12),
    sec.axis = dup_axis(name = NULL)
  ) +
  
  labs(
    title = "S&P Margin",
    subtitle = "Earnings last 12m / sales last 12m, 2std dev band",
    x = NULL,
    y = NULL
  ) +
  coord_cartesian(ylim = c(0, NA))
```

# Model

## Earnings growth


```{r fig.width=9}
bind_rows(
eps  = daily2month(eps_ttm) %>% select(-nobs),
epsf = daily2month(eps_ftm) %>% select(-nobs),
.id = 'key') %>%
  pivot_wider(names_from = 'key') %>% 
  left_join(gdp_monthly, by = 'index') %>%
  mutate(earnings_growth = epsf / eps -1,
         past_growth = (eps / lag(eps,12))^(1/1) - 1,
         gdp_growth = value / lag(value,12) -1) %>% 
  select(index, past_growth, earnings_growth) %>% 
  pivot_longer(-index) %>% 
  ggplot(aes(x=as_date(index), y=value, color = name)) + 
  geom_line(size = 1) + 
  geom_smooth(method = MASS::rlm, formula = y ~ x, se = FALSE, size = 0.7, lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 15),
    sec.axis = dup_axis(name = NULL)
  ) +
  labs(
    title = "S&P Earnings Growth",
    subtitle = "expected growth one year ahead vs growth in past 12m",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme(legend.position = 'bottom')
```


## Earnings next year always up

The chart below shows that forecast earnings are almost always higher than observed earnings in the last 12 months. 

```{r fig.width=9}
bind_rows(
eps  = daily2month(eps_ttm) %>% select(-nobs),
epsf = daily2month(eps_ftm) %>% select(-nobs),
.id = 'key') %>%
  pivot_wider(names_from = 'key') %>% 
  as_tsibble() %>% 
  model(lm = TSLM(log(epsf) ~ log(eps))) %>% 
  augment() %>% 
  add_column(eps = daily2month(eps_ttm) %>% pull(value)) %>% 
  ggplot(aes(x=(eps), y=(epsf))) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  geom_line(aes(y = .fitted), color = 'blue', size = 1) +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +

  scale_x_continuous(
    breaks = scales::breaks_extended(n=15),
    trans = 'log'
    ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 15),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
    ) +
  labs(
    title = "S&P Earnings",
    subtitle = "expected earnings one year ahead - always up",
    x = "EPS last 12 months",
    y = "EPS next 12 months"
  ) 
```


## Drawdown


```{r fig.width=9}
bind_rows(eps = eps_ttm, epsf = eps_ftm, div = spx_div, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  mutate(earnings = eps * div,
         fwdearnings = epsf * div) %>% 
  drop_na(earnings) %>% 
  mutate(
    max_earnings = cummax(earnings),
    gap_earnings = earnings / max_earnings - 1,
    max_fwdearnings = cummax(fwdearnings),
    gap_fwdearnings = fwdearnings / max_fwdearnings - 1
    ) %>% 
  ggplot(aes(x=index, y=gap_earnings)) + 
  geom_line() + 
  geom_area(alpha = 0.2, fill = 'red') + 
  geom_line(aes(x=index, y=gap_fwdearnings), color = 'blue')+
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 15),
    sec.axis = dup_axis(name = NULL)
  ) +
  labs(
    title = "S&P Earnings Drawdown",
    subtitle = "trailing 12m earnings (fwd earnings in blue)",
    x = NULL,
    y = '% from previous high'
  ) + 
  coord_cartesian(ylim = c(NA,0), expand = FALSE)
```


```{r fig.width=9}
spx_price %>%  
  drop_na() %>% 
  mutate(max_price = cummax(value),
         gap_price = value / max_price - 1) %>% 
  ggplot(aes(x=index, y=gap_price)) + 
  geom_line() + 
  geom_area(alpha = 0.2, fill = 'red') + 
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 15),
    sec.axis = dup_axis(name = NULL)
  ) +
  labs(
    title = "S&P Drawdown",
    subtitle = "price index",
    x = NULL,
    y = '% from previous high'
  ) + 
  coord_cartesian(ylim = c(NA,0), expand = FALSE)
```



```{r}
# PE chart
bind_rows(eps = eps_ttm, price = spx_price, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  drop_na() %>% 
  mutate(pe = price / eps) %>% 
  as_tsibble() %>% 
  model(lm = TSLM(log(pe) ~trend())) %>% 
  augment() %>% 
  ggplot(aes(x=index, y=pe)) + 
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +

  labs(
    title = "S&P Price / Earnings ratio",
    subtitle = "Trailing 12-months, 2std dev band",
    x = NULL,
    y = NULL
  )


  


```


```{r}
# PE fowd chart
bind_rows(eps = eps_ftm, price = spx_price, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  drop_na() %>% 
  mutate(pe = price / eps) %>% 
  as_tsibble() %>% 
  model(lm = TSLM(log(pe) ~trend())) %>% 
  augment() %>% 
  ggplot(aes(x=index, y=pe)) + 
  geom_line(size = 1) +
  geom_line(aes(y = .fitted), color = 'blue') +
  geom_line(aes(y = exp(log(.fitted) + 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  geom_line(aes(y = exp(log(.fitted) - 2*sd(.innov))), color = 'blue', lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +

  labs(
    title = "S&P Price / Earnings ratio",
    subtitle = "Next 12-months, 2std dev band",
    x = NULL,
    y = NULL
  )

```



Forward PE is always a bit below trailing PE -- which is another way of saying that expected earnings in the next 12 months are always above trailing 12 months earnings. 

```{r fig.width=8}
bind_rows(
  `trailing earnings` = eps_ttm,
  `forward earnings`  = eps_ftm, .id = 'key')  %>% 
  drop_na() %>% 
  left_join(spx_price, by = 'index') %>% 
  mutate(pe = value.y / value.x) %>% 
  ggplot(aes(x=index, y=pe, color = key)) + 
  geom_line(size = 1) +
  geom_smooth(method = MASS::rlm, formula = y ~ x, se = FALSE, lty = 'dashed') +
  scale_x_date(
    breaks = scales::breaks_pretty(n=20)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    trans = 'log',
    sec.axis = dup_axis(name = NULL)
  ) +

  labs(
    title = "S&P Price / Earnings ratio",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme(legend.position = 'bottom')


```



## Model for future earnings

### Nominal GDP

```{r message=FALSE, warning=FALSE}
gdp_nominal %>% 
  mutate(yoy = value / lag(value,4) - 1) %>% 
  drop_na() %>% 
  ggplot(aes(x = as_date(index), y = yoy)) +
  geom_line(size = 1) +
  geom_smooth(se = FALSE) +
  scale_x_date(
    breaks = scales::breaks_pretty(n = 12)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 15),
    labels = scales::label_percent(accuracy = 0.1),
    sec.axis = dup_axis()
  ) +
  labs(
    title = "US Nominal GDP growth",
    x = NULL,
    y = NULL
  )
  
```

```{r}
# gdp_nominal %>% as_tsibble(index = index) %>% 
#   drop_na() %>% 
#   #rolling 20 years
#   slide_tsibble(.size = 20*4) %>% 
#   model(
#     #aan  =  ETS(log(value) ~ error("A") + trend("A") + season("N")),
#     #aadn =  ETS(log(value) ~ error("A") + trend("Ad") + season("N")),
#     #man  =  ETS(log(value) ~ error("M") + trend("A") + season("N")),
#     #madn =  ETS(log(value) ~ error("M") + trend("Ad") + season("N")),
#     ets =  ETS(log(value)),
#     arima = ARIMA(log(value))
#   ) -> fit
# 
# fc <- fit %>%
#   forecast(h = '2 years')
# 
# fc %>% accuracy(gdp_nominal %>% as_tsibble() %>% 
#   drop_na())
# 
# fc %>% 
#   accuracy(
#     gdp_nominal %>% as_tsibble(),
#     measures = list(interval_accuracy_measures, distribution_accuracy_measures)
#   )
# 
# fc %>% 
#   accuracy(
#     gdp_nominal %>% as_tsibble(),
#     list(skill = skill_score(CRPS))
#   )

# Best model ends up being aan
gdp_nominal %>% as_tsibble(index = index) %>% 
  drop_na() %>% 
  filter_index("2005 Q1" ~ "2019 Q4") %>%
  model(
    aan  =  ETS(log(value)) # ~ error("A") + trend("A") + season("N"))
  ) -> fit

fc <- fit %>%
  forecast(h = '4 years')


fc %>%
  autoplot(gdp_nominal %>% as_tsibble(index = index) %>% drop_na() %>% filter_index("2000 Q1" ~.)) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit)) +
  labs(y="$ billion", title="US GDP nominal") +
  guides(colour = "none") +
  scale_y_continuous(
    labels = scales::label_dollar(accuracy = 1),
    trans = 'log',
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis(name = NULL)
  )



# fit %>%
#   generate(h = 8, times = 5,
#            bootstrap_block_size = 8) %>%
#   autoplot(.sim) +
#   autolayer(gdp_nominal %>% as_tsibble() %>% drop_na() %>% filter_index("2010 Q1" ~.), value) +
#   guides(colour = "none") +
#   scale_y_continuous(
#     labels = scales::label_dollar(accuracy = 1),
#     trans = 'log',
#     breaks = scales::breaks_extended(n = 10)
#   )


```
The table below shows the implied nominal growth in GDP for the 80% forecast range.

```{r}
last_gdp <- gdp_nominal %>% drop_na() %>% tail(1) %>% pull(value)

hilo(fc, 80) %>% tail(1) %>% 
  mutate(p10 = (quantile(value, p = 0.1) / last_gdp - 1)*100,
         p90 = (quantile(value, p = 0.9) / last_gdp - 1)*100,
         .mean = (.mean / last_gdp - 1)*100) %>%
  select(index, p10, .mean, p90)

```

### Sales to GDP

```{r}
db_sales <- bind_rows(sps = sales_ttm, div = spx_div, .id = 'key') %>% 
  pivot_wider(names_from = 'key', values_from = value) %>%
  transmute(index, value = sps * div) %>% 
  drop_na() %>% 
  daily2month() %>% 
  left_join(gdp_monthly, by = 'index', suffix = c('.sales', '.gdp')) %>% 
  mutate(sales_share = value.sales * 10^(-6) / value.gdp) %>% 
  drop_na(sales_share) %>% 
  as_tsibble(index = index) %>% 
  drop_na() 


# db_sales %>% 
#    #rolling 20 years
#    slide_tsibble(.size = 20*4) %>% 
#    model(
#      aan  =  ETS(sales_share ~ error("A") + trend("A") + season("N")),
#      aadn =  ETS(sales_share ~ error("A") + trend("Ad") + season("N")),
#      man  =  ETS(sales_share ~ error("M") + trend("A") + season("N")),
#      madn =  ETS(sales_share ~ error("M") + trend("Ad") + season("N")),
#      ets =  ETS(sales_share),
#      arima = ARIMA(sales_share)
#    ) -> fit_sales
  
  
# fc_sales <- fit_sales %>%
#   forecast(h = '2 years')
# 
# fc_sales %>% accuracy(db_sales) %>% arrange(MAPE)
  
# BEst model is AADn
db_sales %>% 
  filter_index("2005 Q1" ~ "2019 Q4") %>%
  model(
    aadn  =  ETS(sales_share ~ error("A") + trend("Ad") + season("N"))
  ) -> fit_sales

fc_sales <- fit_sales %>%
  forecast(h = '4 years')


fc_sales %>% 
  autoplot(db_sales %>% filter_index("2000 Q1" ~.)) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit_sales)) +
  labs(title="Sales to GDP ratio") +
  guides(colour = "none") +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis(name = NULL)
  )





  #model(lm = TSLM(sales_share ~trend()))
```

```{r}
hilo(fc_sales, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(sales_share, p = 0.2),
         p80 = quantile(sales_share, p = 0.8)) %>%
  select(index, p20, .mean, p80)
```

### Margin

```{r}
db_margin <- bind_rows(sps = sales_ttm, eps = eps_ttm, .id = 'key') %>% 
  pivot_wider(names_from = 'key', values_from = value) %>%
  transmute(index, value = eps / sps) %>% 
  drop_na() %>% 
  daily2month() %>% 
  as_tsibble(index = index) 

 # db_margin %>%
 #    #rolling 20 years
 #    slide_tsibble(.size = 20*4) %>%
 #    model(
 #      aan  =  ETS(value ~ error("A") + trend("A") + season("N")),
 #      aadn =  ETS(value ~ error("A") + trend("Ad") + season("N")),
 #      man  =  ETS(value ~ error("M") + trend("A") + season("N")),
 #      madn =  ETS(value ~ error("M") + trend("Ad") + season("N")),
 #      ets =  ETS(value),
 #      arima = ARIMA(value)
 #    ) -> fit_margin
 # 
 # 
 # fc_margin <- fit_margin %>%
 #   forecast(h = '2 years')
 # 
 # fc_margin %>% accuracy(db_margin) %>% arrange(MAPE)

#AAdn eh o melhor
db_margin %>% 
  filter_index("2005 Q1" ~ "2019 Q4") %>%
  model(
    aadn  =  ETS(value ~ error("A") + trend("Ad") + season("N"))
  ) -> fit_margin

fc_margin <- fit_margin %>%
  forecast(h = '4 years')


fc_margin %>% 
  autoplot(db_margin %>% filter_index("2000 Q1" ~.)) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit_margin)) +
  labs(title="S&P Margin") +
  guides(colour = "none") +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis(name = NULL)
  )



```

```{r}
hilo(fc_margin, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(value, p = 0.2),
         p80 = quantile(value, p = 0.8)) %>%
  select(index, p20, .mean, p80)

```

### PE

```{r}
db_pe <- bind_rows(eps = eps_ftm, price = spx_price, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  drop_na() %>% 
  transmute(index, value = price / eps) %>% 
  daily2month() %>%
  as_tsibble(index = index) 


 # db_pe %>%
 #    #rolling 20 years
 #    slide_tsibble(.size = 20*4) %>%
 #    model(
 #      aan  =  ETS(value ~ error("A") + trend("A") + season("N")),
 #      aadn =  ETS(value ~ error("A") + trend("Ad") + season("N")),
 #      man  =  ETS(value ~ error("M") + trend("A") + season("N")),
 #      madn =  ETS(value ~ error("M") + trend("Ad") + season("N")),
 #      ets =  ETS(value),
 #      arima = ARIMA(value)
 #    ) -> fit_pe
 # 
 # 
 # fc_pe <- fit_pe %>%
 #   forecast(h = '2 years')
 # 
 # fc_pe %>% accuracy(db_pe) %>% arrange(MAPE)
 # 
 #MADN

db_pe %>% 
  filter_index("2005 Q1" ~ "2019 Q4") %>%
  model(
    madn  =  ETS(value ~ error("M") + trend("Ad") + season("N"))
  ) -> fit_pe

fc_pe <- fit_pe %>%
  forecast(h = '4 years')


fc_pe %>% 
  autoplot(db_pe %>% filter_index("2000 Q1" ~.)) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit_pe)) +
  labs(title="S&P Fwd PE") +
  guides(colour = "none") +
  scale_y_continuous(
   # labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis(name = NULL)
  )


```

```{r}
hilo(fc_pe, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(value, p = 0.2),
         p80 = quantile(value, p = 0.8)) %>%
  select(index, p20, .mean, p80)
```


### Assumptions

```{r}
h_gdp <- hilo(fc, 80) %>% tail(1) %>% 
  mutate(p10 = quantile(value, p = 0.2),
         p90 = quantile(value, p = 0.8)) %>%
  as_tibble() %>%
  select(p10, .mean, p90) 

h_sales <- hilo(fc_sales, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(sales_share, p = 0.2),
         p80 = quantile(sales_share, p = 0.8)) %>%
  as_tibble() %>%
  select(p20, .mean, p80)

h_margin <- hilo(fc_margin, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(value, p = 0.2),
         p80 = quantile(value, p = 0.8)) %>%
  as_tibble() %>%
  select(p20, .mean, p80)

h_pe <- hilo(fc_pe, 80) %>% tail(1) %>% 
  mutate(p20 = quantile(value, p = 0.2),
         p80 = quantile(value, p = 0.8)) %>%
  as_tibble() %>%
  select(p20, .mean, p80)


tibble(gdp = t(h_gdp), sales = t(h_sales), margin = t(h_margin), pe = t(h_pe)) %>%
  expand(gdp, sales, margin, pe) %>%
  mutate(earnings = gdp * 10^6 * sales * margin,
         eps = earnings / 8455447,
         spx = pe * eps) %>%
  ggplot(aes(x = spx)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 15)+
 geom_density(alpha=.2, fill="#FF6666") 
  
```

```{r paged.print=TRUE}
tibble(gdp = t(h_gdp), sales = t(h_sales), margin = t(h_margin), pe = t(h_pe)) %>%
  expand(gdp, sales, margin, pe) %>%
  mutate(earnings = gdp * 10^6 * sales * margin,
         eps = earnings / 8455447,
         spx = pe * eps) %>% arrange(desc(eps)) %>%
  select(-earnings)
```

## PE vs real rates

```{r}
db_realrate <- bind_rows(eps = eps_ftm, price = spx_price, .id = 'key') %>% 
  pivot_wider(names_from = 'key') %>% 
  drop_na() %>% 
  transmute(index, value = price / eps) %>% 
  left_join(us10y_real, by = 'index', suffix = c('.pe', '.real10y')) %>% 
  drop_na() %>% 
  rename(pe = value.pe, real10y = value.real10y) %>% 
  as_tsibble(index = index)


db_realrate %>% 
  filter_index("2015-01-01" ~ .) %>% 
  #mutate(regime = if_else(index >= "2012-01-01", 1, 0)) %>% 
  model(
    lm = TSLM(pe ~ real10y)
  ) -> fit_rr
  
report(fit_rr)  

augment(fit_rr) %>%
  ggplot(aes(x = index)) +
  geom_line(aes(y = pe, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
    title = "SPX P/E next 12 months"
  ) +
  scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) +
  guides(colour = guide_legend(title = NULL))

#fit_rr %>% gg_tsresiduals()

```

