---
title: "US Government Debt"
author: "Dr. Paulo Gustavo Grahl, CFA"
date: "2022-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE,
    include = TRUE
)

library(tidyverse)
library(tsibble)
library(lubridate)
library(patchwork)
```

```{r}
source("code/download_public_debt.R", local = knitr::knit_global())
source("code/fed_soma_holdings.R", local = knitr::knit_global())

# functions
my_dollar_bn <- function(x) scales::label_dollar(accuracy = 1, scale = 1e-3)(x)
```



## Introduction

The **total public debt** is at `r my_dollar_bn(tail(debt_total,1)$value)`bn (`r format(tail(debt_total,1)$index, "%b %d")`), and **debt held by the public** reached `r my_dollar_bn(tail(debt_public,1)$value)`bn (`r format(tail(debt_public,1)$index, "%b %d")`). The Federal Reserve holds `r my_dollar_bn(tail(treasury_holdings,1)$value)`bn in its balance sheet (`r format(tail(treasury_holdings,1)$index, "%b %d")`), or `r round((tail(treasury_holdings,1)$value/filter(debt_public, index == tail(treasury_holdings,1)$index)$value)*100,1)`% of the total debt held by the public.

```{r fig.width=8}
bind_rows(treasury_holdings, debt_public) %>% 
  filter(index >= "2006-01-01") %>% 
  ggplot(aes(x = index, y = value, color = key)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-3),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "US Public Debt",
    subtitle = "US$ billions",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank()
        )
```

```{r fig.width=8}
left_join(treasury_holdings, debt_public, by = "index") %>% 
  drop_na() %>% 
  transmute(index, key = "Treasury Holdings / Debt Held by the Public", 
            value = value.x / value.y) %>%
  ggplot(aes(x = index, y = value, color = key)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = scales::breaks_width(0.02),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "Fed Holdings of Treasury",
    subtitle = "% of debt held by the public",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank())
```

```{r}
# convert to 12 month absolute change
db12m <-
  bind_rows(treasury_holdings, debt_public) %>% 
  pivot_wider(names_from = key, values_from = value) %>% 
  arrange(index) %>% 
  drop_na() %>% 
  pivot_longer(-index) %>% 
  as_tsibble(index = index, key = name) %>%   
  index_by(month = ~yearmonth(.)) %>% 
  group_by_key() %>% 
  summarise(value = last(value)) %>% 
  group_by_key() %>% 
  mutate(value = value - lag(value, 12)) %>% 
  drop_na() %>% 
  ungroup()
```

Another interesting fact is that total public **debt held by the public** increased `r my_dollar_bn(filter(db12m, name == "Debt Held by the Public") %>% slice_tail() %>% pull(value))`bn, while **Federal Reserve holdings** rose by `r my_dollar_bn(filter(db12m, name == "Fed Treasury Holdings") %>% slice_tail() %>% pull(value))`bn.

```{r fig.width=8}
db12m %>% 
  filter_index("2007 Jan" ~. ) %>% 
  ggplot(aes(x = as.Date(month), y = value, color = name)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-3),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "US Public Debt",
    subtitle = "Change in 12 months, US$ billions",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank()
        )
  
```

The chart below show the amout of Treasury debt that was **absorbed by the public, excluding Fed**.

```{r fig.width=8}
db12m %>% 
  pivot_wider(names_from = name, values_from = value) %>% 
  mutate(`Held by the Public (ex Fed)` = 
           `Debt Held by the Public` - `Fed Treasury Holdings`) %>% 
  select(month, `Held by the Public (ex Fed)`) %>% 
  pivot_longer(-month) %>% 
  filter_index("2007 Jan" ~. ) %>% 
  ggplot(aes(x = as.Date(month), y = value, color = name)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-3),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "US Public Debt",
    subtitle = "Change in 12 months, US$ billions",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank()
        )
  
```


## Fed Purchases


```{r}
bind_rows(treasury_bills, treasury_bonds, treasury_tips) %>% 
  mutate(key = str_remove_all(key, "Fed|Holdings")) %>% 
  filter(index >= "2020-01-01") %>% 
  ggplot(aes(x = index, y = value, color = key)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-3),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "US Federal Reserve Holdings by type",
    subtitle = "US$ billions",
    x = NULL,
    y = NULL,
    color = "Type of Holdings"
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank()
        )

  
```


```{r}
bind_rows(treasury_bills, treasury_bonds, treasury_tips) %>% 
  mutate(key = str_remove_all(key, "Fed|Holdings")) %>% 
  group_by(key) %>% 
  mutate(change = value - lag(value,4)) %>% 
  ungroup() %>% 
  filter(index >= "2020-01-01") %>% 
  ggplot(aes(x = index, y = change, color = key)) + 
  geom_line(size = 1) +
  annotate('text', x = as.Date("2020-07-01"), y = 95000, 
           label = stringr::str_wrap("Fed purchases reached a bit more than $1tn in 4w during the initial months of Covid", width = 45),
           hjust = 0) + 
  scale_y_continuous(
    labels = scales::label_dollar(scale = 1e-3),
    breaks = scales::breaks_extended(n = 10),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "US Federal Reserve QE - 4 weeks rolling purchases",
    subtitle = "US$ billions (only Treasury assets, excludes MBS, CMBS)",
    x = NULL,
    y = NULL,
    color = "Type of Holdings"
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank()
        ) +
  coord_cartesian(ylim = c(NA, 100000))

```

```{r}
bind_rows(treasury_tips, treasury_holdings) %>% 
  pivot_wider(names_from = key,  values_from = value) %>% 
  mutate(`Tips share` = `Fed Tips Holdings` / `Fed Treasury Holdings`) %>%
  select(index, `Tips share`) %>% 
  pivot_longer(-index) %>% 
  filter(index >= "2014-01-01") %>% 
  ggplot(aes(x = index, y = value, color = name)) + 
  geom_line(size = 1) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1),
    breaks = scales::breaks_width(0.005),
    sec.axis = dup_axis()
  ) +
  scale_x_date(
    labels = scales::label_date_short(),
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "Fed Holdings of Tips",
    subtitle = "% total Fed holdings",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.border = element_blank())
  
```
