---
title: "us-inflation-simulation"
author: "pgg1309"
date: "2022-02-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  include = TRUE
)
require(fpp3)
```

```{r functions}
### Functions ----
add_lags <- function(original, variable_names, lags = c(1, 2)) {
  for (variable_name in variable_names) {
    for (lag in lags) {
      new_variable_name <- paste0("lag", lag, "_", variable_name)
      new_value <- dplyr::lag(dplyr::pull(original, variable_name), lag)
      original <- dplyr::mutate(original, !!new_variable_name := new_value)
    }
  }
  return(original)
}

mygg_tsresiduals <- function(data, type = "innovation", plot_type = "histogram", ...) {
  if (!fabletools::is_mable(data)) {
    abort("gg_tsresiduals() must be used with a mable containing only one model.")
  }
  data <- stats::residuals(data, type = type)
  if (n_keys(data) > 1) {
    abort("gg_tsresiduals() must be used with a mable containing only one model.")
  }
  out <- gg_tsdisplay(data, !!sym(".resid"),
    plot_type = plot_type,
    ...
  )
  out[[1]] <- out[[1]] + ggplot2::ylab(sub("([[:alpha:]])(.+)",
    "\\U\\1\\L\\2 residuals", type,
    perl = TRUE
  ))
  out
}

```

```{r loaddata}
cpiboard <- pins::board_folder("E:/RData")
cpi <- pins::pin_read(cpiboard, "USCPI")
fredr::fredr_set_key(options()$fred.key)

# Energy (commodities and services)
energy <- filter(cpi, series_id == "CUSR0000SA0E") %>%
  select(month, value) %>%
  mutate(key = "cpi_energy") %>%
  update_tsibble(key = "key")

# CPI ex energy
cpix <- filter(cpi, series_id == "CUSR0000SA0LE") %>%
  select(month, value) %>% 
  mutate(key = "cpi_x_energy") %>% 
  update_tsibble(key = "key")

# Core goods CPI (commodities less food and energy)
core_goods <- filter(cpi, series_id == "CUSR0000SACL1E") %>% 
  select(month, value) %>% 
  mutate(key = "core_goods") %>% 
  update_tsibble(key = "key")

# Core services CPI (services less energy services)
core_services <- filter(cpi, series_id == "CUSR0000SASLE") %>% 
  select(month, value) %>% 
  mutate(key = "core_services") %>% 
  update_tsibble(key = 'key')


# Function to read from FRED
get_fred_data <- function(id, keyname) {
  fredr::fredr(
    series_id = id,
    frequency = "d"
  ) %>%
    group_by(month = yearmonth(date)) %>%
    summarize(value = mean(value, na.rm = TRUE)) %>%
    mutate(key = keyname) %>%
    tsibble(index = month, key = "key") %>%
    return()
}

# Brent and WTI
wti <- get_fred_data("DCOILWTICO", "wti")
brent <- get_fred_data("DCOILBRENTEU", "brent")

```

# US CPI - possible scenarios

## Energy CPI

The chart below plots the *z-score* of the CPI energy index, WTI and Brent oil prices. Adjusting for volatility, it is clear that oil prices and the energy CPI index share a similar trend.

```{r energy_chart}
bind_rows(energy, wti, brent) %>%
  # limit sample to Jan 1986, start of WTI data
  # brent sample starts at May 1987
  filter_index("1986 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = as.vector(scale(log(value)))) %>%
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "Energy CPI index vs. oil price",
    x = NULL,
    y = "z-score",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())
```

### A simple model

Despite the clear association between energy CPI and oil prices, a simple model fails to capture the dynamics.

```{r simple_model}
simple_mod <- bind_rows(energy, wti) %>%
  pivot_wider(names_from = key, values_from = value) %>%
  drop_na() %>%
  model(linear = TSLM(log(cpi_energy) ~ log(wti)))

report(simple_mod)
```

```{r simple_model_fit}
augment(simple_mod) %>%
  ggplot(aes(x = month)) +
  geom_line(aes(y = cpi_energy, colour = "Data"), size = 1) +
  geom_line(aes(y = .fitted, colour = "Fitted"), size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "CPI Energy index",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())
```

```{r simple_model_scatter}
augment(simple_mod) %>%
  ggplot(aes(x = cpi_energy, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  labs(
    title = "CPI Energy index",
    x = "Data (actual values)",
    y = "Fitted (predicted values)",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())
```

Looking at the *residuals* it is clear that this simple model is misspecified as there is autocorrelation in the residuals.

```{r simple_model_resid}
simple_mod %>% mygg_tsresiduals(plot_type = 'partial')
```


### Additional models

The residuals of the *simple model* reported in the previous section clearly shows the presence of a *trend* or *drift*. Moreover, even allowing for the presence of a trend the residuals present serial correlation.

Therefore it seems appropriated to model energy cpi considering a trend and assuming an ARMA process for the residuals. 

A *deterministic trend* is represented by

$$ ecpi_t = \beta_0 + \beta_1 wti_t + \beta_2 t + \eta_t$$

where $\eta_t$ is an ARMA process. A *stochastic trend* is represented by

$$ecpi_t = ecpi_{t-1} + \beta_1 wti_t + \beta_2 + \eta_t $$

Additionally, I've tested both models including up to four lags for the $wti_t$ prices.

The table blow shows the different models considered, using data up to December 2019, and sorted by AICc.

```{r models}
# split data into test and training --------------------------------------------
full_data <- bind_rows(energy, wti) %>%
  pivot_wider(names_from = key, values_from = value) %>%
  drop_na(wti) # keeps wti

train_data <- filter_index(full_data, . ~ "2019 Dec")
test_data <- filter_index(full_data, "2020 Jan" ~ .)

# Estimate models --------------------------------------------------------------
table_mod <- train_data %>%
  model(
    deterministic = ARIMA(log(cpi_energy) ~ pdq(d = 0) + PDQ(0, 0, 0) +
      1 + trend() + log(wti),
    greedy = FALSE, approximation = FALSE
    ),
    deterministic_l1 = ARIMA(log(cpi_energy) ~ pdq(d = 0) + PDQ(0, 0, 0) +
      1 + trend() + log(wti) + lag(log(wti)),
    greedy = FALSE, approximation = FALSE
    ),
    deterministic_l2 = ARIMA(log(cpi_energy) ~ pdq(d = 0) + PDQ(0, 0, 0) +
      1 + trend() + log(wti) + lag(log(wti)) +
      lag(log(wti), 2),
    greedy = FALSE, approximation = FALSE
    ),
    deterministic_l3 = ARIMA(log(cpi_energy) ~ pdq(d = 0) + PDQ(0, 0, 0) +
      1 + trend() + log(wti) + lag(log(wti)) +
      lag(log(wti), 2) + lag(log(wti), 3),
    greedy = FALSE, approximation = FALSE
    ),
    deterministic_l4 = ARIMA(log(cpi_energy) ~ pdq(d = 0) + PDQ(0, 0, 0) +
      1 + trend() + log(wti) + lag(log(wti)) +
      lag(log(wti), 2) + lag(log(wti), 3) +
      lag(log(wti), 4),
    greedy = FALSE, approximation = FALSE
    ),
    stochastic = ARIMA(log(cpi_energy) ~ pdq(d = 1) + PDQ(0, 0, 0) +
      1 + log(wti),
    greedy = FALSE, approximation = FALSE
    ),
    stochastic_l1 = ARIMA(log(cpi_energy) ~ pdq(d = 1) + PDQ(0, 0, 0) +
      1 + log(wti) + lag(log(wti)),
    greedy = FALSE, approximation = FALSE
    ),
    stochastic_l2 = ARIMA(log(cpi_energy) ~ pdq(d = 1) + PDQ(0, 0, 0) +
      1 + log(wti) + lag(log(wti)) + lag(log(wti), 2),
    greedy = FALSE, approximation = FALSE
    ),
    stochastic_l3 = ARIMA(log(cpi_energy) ~ pdq(d = 1) + PDQ(0, 0, 0) +
      1 + log(wti) + lag(log(wti)) + lag(log(wti), 2) +
      lag(log(wti), 3),
    greedy = FALSE, approximation = FALSE
    ),
    stochastic_l4 = ARIMA(log(cpi_energy) ~ pdq(d = 1) + PDQ(0, 0, 0) +
      1 + log(wti) + lag(log(wti)) + lag(log(wti), 2) +
      lag(log(wti), 3) + lag(log(wti), 4),
    greedy = FALSE, approximation = FALSE
    )
  )

glance(table_mod) %>% arrange(AICc) 




# select(table_mod, deterministic_l1) %>%
#   forecast(test_data) %>%
#   tail()
# autoplot(filter_index(full_data, "2010 Jan" ~ .))
# 
# select(table_mod, stochastic_l1, deterministic_l1) %>%
#   forecast(test_data) %>%
#   autoplot(filter_index(full_data, "2010 Jan" ~ .))
# 
# 
# 
# 


```

Looking at the AICc information criteria the best model includes a deterministic trend, one lag of $wti$, and ARMA(1,3) errors. The BIC criteria suggest a stochastic trend model with one lag of $wti$ and MA(3) errors. 

Both models are printed below:

```{r selected_models}
table_mod %>%
  select(deterministic_l1) %>%
  report()

table_mod %>%
  select(stochastic_l1) %>%
  report()

```

And the model's fitted values are similar and close to the actual observed values.

```{r}
table_mod %>%
  select(deterministic_l1, stochastic_l1) %>%
  refit(new_data = full_data) %>% 
  augment() %>% 
  ggplot(aes(x = cpi_energy, y = .fitted, color = .model)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  labs(
    title = "CPI Energy index",
    x = "Data (actual values)",
    y = "Fitted (predicted values)",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())
```

And looking at the residuals it seems that both models are appropriate.

Deterministic trend: 

```{r}
table_mod %>%
  select(deterministic_l1) %>%
  mygg_tsresiduals()
```

Stochastic trend:

```{r}
table_mod %>%
  select(stochastic_l1) %>%
  mygg_tsresiduals()
```

## Outlook for Energy CPI

### WTI Forecast 

In order to use the model to forecast the energy component of CPI one needs to assume a path for oil prices. One possibility is to assume a *random walk* model. The chart below shows some possible simulated paths for the next 2 years based on bootstrap residuals.

```{r wti_model}
wti_model <- wti %>% 
  model(
    rw = NAIVE(value),
    drift = NAIVE(value ~ drift())
    )
wti_sim <- select(wti_model, rw) %>% 
  generate(h=24, times = 5, boostrap = TRUE)
```

```{r}
wti %>% 
  ggplot(aes(x = month)) +
  geom_line(aes(y = value), size = 1) +
  geom_line(aes(y = .sim, color = as.factor(.rep)), data = wti_sim, size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "WTI monthly average",
    color = NULL,
    y = "US$ / bbl",
    x = NULL
  ) + 
  guides(color = 'none')
```

However, if we simulate enough paths and take the average we get the the expected *random walk* behavior. 

```{r}
wti_forecast <- select(wti_model, rw) %>% forecast(h = 24, bootstrap = TRUE)
wti_forecast %>% 
  autoplot(select(wti, -key)) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
    labs(
      title = "WTI monthly average",
        y = "US$ / bbl",
    x = NULL
    ) 
```


But we can also pick a quantile if we want to add a *bias* to the path of future oil prices. In the chart below, the path highlighted in red shows the 80^th^ percentile, or in other words, the path where there are only 20% of the sample above it. The selected path can then be used to forecast energy CPI.


```{r}
wti_forecast %>% 
  autoplot(select(wti, -key)) +
  geom_line(
    aes(y = hl),
    color = 'red',
    size = 1,
    data = mutate(
      wti_forecast,
      hl  = hilo(value, 60)$upper)
    ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
    labs(
      title = "WTI monthly average",
      subtitle = "red line is the 80th percentile",
        y = "US$ / bbl",
    x = NULL
    ) 

```

### Energy CPI assuming oil in the 80^th^ percentile

```{r}
wti_future <- wti_forecast %>% 
  mutate(wti = hilo(value, 60)$upper) %>% 
  pull(wti)

future_data <- new_data(full_data, 24) %>% 
  mutate(wti = wti_future)

db <- bind_rows(full_data, future_data) %>% 
  group_split(is.na(cpi_energy), .keep = FALSE)


table_mod %>%
  select(stochastic_l1) %>%
  refit(new_data = db[[1]]) %>% 
  forecast(new_data = db[[2]]) %>% 
  autoplot(filter_index(full_data, "2000 Jan" ~.)) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "CPI Energy index",
    color = NULL,
    y = "index",
    x = NULL
  )





```


The expected path for the annual energy inflation would gradually slowdown, but would still be above 15% yoy in the early part of 3Q 2022.
```{r}
full_data %>%
  append_row(24) %>%
  left_join(
    table_mod %>%
      select(stochastic_l1) %>%
      refit(new_data = db[[1]]) %>%
      forecast(new_data = db[[2]]) %>%
      as_tibble() %>%
      select(month, .mean) %>%
      rename(cpi_e_fcast = .mean),
    by = 'month'
  ) %>% 
  mutate(cpi_e_full = if_else(is.na(cpi_energy), cpi_e_fcast, cpi_energy),
         yoy = cpi_e_full / lag(cpi_e_full, 12) - 1,
         forecast = if_else(is.na(cpi_energy), TRUE, FALSE)) %>% 
  filter_index("2016 Jan" ~. ) %>% 
  ggplot(aes(x = month, y = yoy, color = forecast)) +
  geom_line(size = 1) +
  geom_vline(
    xintercept = as_date(yearmonth("2022 Feb")),
    lty = 'dashed') + 
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    labels = scales::label_percent()
  ) + 
   scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "Energy CPI",
    x = NULL,
    y = "% yoy"
  ) +
  theme(legend.position = "bottom")
  

```

### Energy CPI assuming oil in the 40^th^percentile

```{r}
wti_forecast %>% 
  autoplot(select(wti, -key)) +
  geom_line(
    aes(y = hl),
    color = 'red',
    size = 1,
    data = mutate(
      wti_forecast,
      hl  = hilo(value, 20)$lower)
    ) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
    labs(
      title = "WTI monthly average",
      subtitle = "red line is the 40th percentile",
        y = "US$ / bbl",
    x = NULL
    ) 




wti_future <- wti_forecast %>% 
  mutate(wti = hilo(value, 20)$lower) %>% 
  pull(wti)

future_data <- new_data(full_data, 24) %>% 
  mutate(wti = wti_future)

db <- bind_rows(full_data, future_data) %>% 
  group_split(is.na(cpi_energy), .keep = FALSE)


table_mod %>%
  select(stochastic_l1) %>%
  refit(new_data = db[[1]]) %>% 
  forecast(new_data = db[[2]]) %>% 
  autoplot(filter_index(full_data, "2000 Jan" ~.)) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 10)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "CPI Energy index",
    color = NULL,
    y = "index",
    x = NULL
  )

full_data %>%
  append_row(24) %>%
  left_join(
    table_mod %>%
      select(stochastic_l1) %>%
      refit(new_data = db[[1]]) %>%
      forecast(new_data = db[[2]]) %>%
      as_tibble() %>%
      select(month, .mean) %>%
      rename(cpi_e_fcast = .mean),
    by = 'month'
  ) %>% 
  mutate(cpi_e_full = if_else(is.na(cpi_energy), cpi_e_fcast, cpi_energy),
         yoy = cpi_e_full / lag(cpi_e_full, 12) - 1,
         forecast = if_else(is.na(cpi_energy), TRUE, FALSE)) %>% 
  filter_index("2016 Jan" ~. ) %>% 
  ggplot(aes(x = month, y = yoy, color = forecast)) +
  geom_line(size = 1) +
  geom_vline(
    xintercept = as_date(yearmonth("2022 Feb")),
    lty = 'dashed') + 
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    labels = scales::label_percent()
  ) + 
   scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 12),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "Energy CPI",
    x = NULL,
    y = "% yoy"
  ) +
  theme(legend.position = "bottom")


```


## Non-energy CPI
```{r}
bind_rows(energy, cpix) %>%
  # limit sample to Jan 1986, start of WTI data
  # brent sample starts at May 1987
  filter_index("1986 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = as.vector(scale(log(value)))) %>%
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "CPI excluding energy vs. energy cpi",
    x = NULL,
    y = "z-score",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())


# YoY chart -------------------------------------------------------------------

bind_rows(energy, cpix) %>%
  # limit sample to Jan 1986, start of WTI data
  # brent sample starts at May 1987
  filter_index("1985 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = value / lag(value,12) - 1) %>%
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    labels = scales::label_percent()
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "CPI excluding energy vs. energy cpi",
    x = NULL,
    y = "% YoY",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())





```


The chart below shows that non-energy inflation is well above the recent historical ranges, while the energy inflation is in line with past episodes.

```{r}
# Level chart -----------------------------------------------------------------
bind_rows(core_goods, core_services) %>%
  filter_index("1986 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = as.vector(scale(log(value)))) %>%
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12)
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "Core CPI goods vs. services",
    x = NULL,
    y = "z-score",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())



# YoY chart -------------------------------------------------------------------
bind_rows(energy, cpix) %>%
  # limit sample to Jan 1986, start of WTI data
  # brent sample starts at May 1987
  filter_index("1985 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = value / lag(value,12) - 1) %>%
  drop_na() %>% 
  mutate(value = as.vector(scale(value))) %>% 
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15)
  ) +
  labs(
    title = "CPI excluding energy vs. energy cpi",
    x = NULL,
    y = "z-score",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())

```

## Core goods vs core services CPI

```{r}
bind_rows(core_goods, core_services) %>%
  filter_index("1985 Jan" ~ .) %>%
  group_by_key() %>%
  mutate(value = value / lag(value,12) - 1) %>%
  drop_na() %>% 
  ggplot(aes(x = month, y = value, color = key)) +
  geom_line(size = 1) +
  scale_y_continuous(
    breaks = scales::breaks_extended(n = 12),
    labels = scales::label_percent()
  ) +
  scale_x_yearmonth(
    breaks = scales::breaks_pretty(n = 15),
    labels = scales::label_date_short()
  ) +
  labs(
    title = "Core CPI goods vs. services",
    x = NULL,
    y = "% YoY",
    color = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom", panel.border = element_blank())


```




















<!-- ## Scatter plot -->

<!-- The relationship between energy CPI and WTI prices is clearly nonlinear.  -->

<!-- ```{r, fig.asp=1} -->

<!-- bind_rows(energy, wti) %>%  -->

<!--   # limit sample to Jan 1986, start of WTI data -->

<!--   # brent sample starts at May 1987 -->

<!--   filter_index("1986 Jan" ~. ) %>%  -->

<!--   group_by_key() %>%  -->

<!--   mutate(value = as.vector(scale(log(value)))) %>%  -->

<!--   pivot_wider(names_from = key, values_from = value) %>%  -->

<!--   ggplot(aes(x = wti, y = cpi_energy)) + -->

<!--   geom_point() +  -->

<!--   geom_hline(yintercept = 0, lty = 'dashed') + -->

<!--   geom_vline(xintercept = 0, lty = 'dashed') + -->

<!--   geom_abline(slope = 1, intercept = 0, lty = 'dashed') + -->

<!--   geom_smooth(se = TRUE) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--     ) + -->

<!--   labs( -->

<!--     title = "Energy CPI index vs. oil price", -->

<!--     x = 'WTI (z-score)', -->

<!--     y = 'CPI energy (z-score)', -->

<!--     color = NULL -->

<!--     ) + -->

<!--   theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- Looking at the YoY changes -->

<!-- ```{r} -->

<!-- bind_rows(energy, wti) %>%  -->

<!--   # limit sample to Jan 1986, start of WTI data -->

<!--   # brent sample starts at May 1987 -->

<!--   group_by_key() %>%  -->

<!--   mutate(value = log(value/lag(value,12))) %>%  -->

<!--   filter_index("1986 Jan" ~. ) %>%  -->

<!--   pivot_wider(names_from = key, values_from = value) %>%  -->

<!--   drop_na() %>%  -->

<!--   ggplot(aes(x = wti, y = cpi_energy)) + -->

<!--   geom_point() +  -->

<!--   geom_hline(yintercept = 0, lty = 'dashed') + -->

<!--   geom_vline(xintercept = 0, lty = 'dashed') + -->

<!--   geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, bs = "cs")) +  -->

<!--   geom_smooth(se = FALSE, method = MASS::rlm, lty = 'dashed', formula = y ~ x) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12), -->

<!--     labels = scales::label_percent() -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12), -->

<!--     labels = scales::label_percent() -->

<!--     ) + -->

<!--   labs( -->

<!--     title = "Energy CPI index vs. oil price", -->

<!--     x = 'WTI (yoy)', -->

<!--     y = 'CPI energy (yoy)', -->

<!--     color = NULL -->

<!--     ) + -->

<!--   theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- Looking at quarterly changes -->

<!-- ```{r} -->

<!-- bind_rows(energy, wti) %>%  -->

<!--   # limit sample to Jan 1986, start of WTI data -->

<!--   # brent sample starts at May 1987 -->

<!--   group_by_key() %>%  -->

<!--   mutate(value = log(value/lag(value,3))) %>%  -->

<!--   filter_index("1986 Jan" ~. ) %>%  -->

<!--   pivot_wider(names_from = key, values_from = value) %>%  -->

<!--   drop_na() %>%  -->

<!--   ggplot(aes(x = wti, y = cpi_energy)) + -->

<!--   geom_point() +  -->

<!--   geom_hline(yintercept = 0, lty = 'dashed') + -->

<!--   geom_vline(xintercept = 0, lty = 'dashed') + -->

<!--   geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, bs = "cs")) +  -->

<!--   geom_smooth(se = FALSE, method = MASS::rlm, lty = 'dashed', formula = y ~ x) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12), -->

<!--     labels = scales::label_percent() -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12), -->

<!--     labels = scales::label_percent() -->

<!--     ) + -->

<!--   labs( -->

<!--     title = "Energy CPI index vs. oil price", -->

<!--     x = 'WTI (m/m-3)', -->

<!--     y = 'CPI energy (m/m-3)', -->

<!--     color = NULL -->

<!--     ) + -->

<!--   theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ## Addint a trend  -->

<!-- ```{r} -->

<!-- trend_mod <- bind_rows(energy, wti) %>%  -->

<!--   pivot_wider(names_from = key, values_from = value) %>%  -->

<!--   drop_na() %>%  -->

<!--   model(trend = TSLM(log(cpi_energy) ~ log(wti) + trend()))  -->

<!-- report(trend_mod) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(trend_mod) %>%  -->

<!--   ggplot(aes(x = month)) +  -->

<!--   geom_line(aes(y = cpi_energy, colour = "Data"), size = 1) + -->

<!--   geom_line(aes(y = .fitted, colour = "Fitted"), size = 1) + -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_yearmonth( -->

<!--     breaks = scales::breaks_pretty( n = 15), -->

<!--     labels = scales::label_date_short() -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = NULL, -->

<!--     y = NULL, -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(trend_mod) %>%  -->

<!--   ggplot(aes(x = cpi_energy, y = .fitted)) +  -->

<!--   geom_point() + -->

<!--   geom_abline(intercept = 0, slope = 1) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = "Data (actual values)", -->

<!--     y = "Fitted (predicted values)", -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- trend_mod %>% mygg_tsresiduals() -->

<!-- trend_mod %>% mygg_tsresiduals(plot_type = 'partial') -->

<!-- ``` -->

<!-- report(ar_mod) -->

<!-- glance(ar_mod) %>% select(.model, adj_r_squared, CV, AIC, AICc, BIC) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(ar_mod) %>%  -->

<!--   ggplot(aes(x = month)) +  -->

<!--   geom_line(aes(y = cpi_energy, colour = "Data"), size = 1) + -->

<!--   geom_line(aes(y = .fitted, colour = "Fitted"), size = 1) + -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_yearmonth( -->

<!--     breaks = scales::breaks_pretty( n = 15), -->

<!--     labels = scales::label_date_short() -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = NULL, -->

<!--     y = NULL, -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(ar_mod) %>%  -->

<!--   ggplot(aes(x = cpi_energy, y = .fitted, color = .model)) +  -->

<!--   geom_point() + -->

<!--   geom_abline(intercept = 0, slope = 1) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = "Data (actual values)", -->

<!--     y = "Fitted (predicted values)", -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- ar_mod %>% select(ar0) %>% mygg_tsresiduals() -->

<!-- ar_mod %>% select(ar5) %>% mygg_tsresiduals(plot_type = 'partial') -->

<!-- ar_mod %>% select(ar6) %>% report() -->

<!-- ``` -->

<!-- ## ARIMA -->

<!-- ```{r} -->

<!-- arima_mod <- bind_rows(energy, wti) %>%  -->

<!--   pivot_wider(names_from = key, values_from = value) %>%  -->

<!--   drop_na() %>%  -->

<!--   model( -->

<!--     arima = ARIMA(log(cpi_energy) ~ trend() + log(wti)) -->

<!--   ) -->

<!-- report(arima_mod) -->

<!-- glance(ar_mod) %>% select(.model, adj_r_squared, CV, AIC, AICc, BIC) -->

<!-- arima_mod %>% mygg_tsresiduals(plot_type = 'partial') -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(arima_mod) %>%  -->

<!--   ggplot(aes(x = cpi_energy, y = .fitted)) +  -->

<!--   geom_point() + -->

<!--   geom_abline(intercept = 0, slope = 1) +  -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = "Data (actual values)", -->

<!--     y = "Fitted (predicted values)", -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- augment(arima_mod) %>%  -->

<!--   ggplot(aes(x = month)) +  -->

<!--   geom_line(aes(y = cpi_energy, colour = "Data"), size = 1) + -->

<!--   geom_line(aes(y = .fitted, colour = "Fitted"), size = 1) + -->

<!--   scale_y_continuous( -->

<!--     breaks = scales::breaks_extended(n = 12) -->

<!--   ) + -->

<!--   scale_x_yearmonth( -->

<!--     breaks = scales::breaks_pretty( n = 15), -->

<!--     labels = scales::label_date_short() -->

<!--   ) + -->

<!--   labs( -->

<!--     title = "CPI Energy index", -->

<!--     x = NULL, -->

<!--     y = NULL, -->

<!--     color = NULL -->

<!--   ) + -->

<!--    theme_bw() +  -->

<!--   theme(legend.position = 'bottom', panel.border = element_blank()) -->

<!-- ``` -->

<!-- # A  -->
