---
title: "us-retail-sales"
author: "pgg1309"
date: "2022-05-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE,
    include = TRUE
)
```

```{r}
library(tidyverse)
library(tsibble)
library(patchwork)
library(pls)
#library(fpp3)
#require(tidymodels)
```

```{r local_functions}
#############################################
# Function to extract tables from txt files #
#############################################
extract_tables <- function(txt) {
  # find tables
  xcategories <- grep("^CATEGORIES$", txt)
  xdatatypes  <- grep("DATA TYPES", txt)
  xerrortypes <- grep("ERROR TYPES", txt)
  xgeolevels <- grep("GEO LEVELS", txt)
  xtimeperiods <- grep("TIME PERIODS", txt)
  xnotes <- grep("NOTES", txt)
  xupdated <- grep("DATA UPDATED ON", txt)
  xdata <- grep("^DATA$", txt)
  
  
  # CATEGORY
  categories_names <-
    txt[(xcategories + 1)] %>% str_split_fixed(",", 4) %>% as.vector()
  
  categories_tbl <-
    tibble(
      x1 = txt[(xcategories + 2):(xdatatypes - 1)] %>%
        str_extract("^[0-9]{1,2},") %>% str_remove(","),
      x2 = txt[(xcategories + 2):(xdatatypes - 1)] %>%
        str_extract(",[[:alnum:]]+,") %>% str_remove_all(","),
      x3 = txt[(xcategories + 2):(xdatatypes - 1)] %>%
        str_extract(',\"[[:print:]]+\",') %>%
        str_remove(',\"') %>%
        str_remove('\",'),
      x4 = txt[(xcategories + 2):(xdatatypes - 1)] %>%
        str_extract(",[0-9]$") %>% str_remove(",")
    ) %>%
    rename_with(~ categories_names)
  
  # TIME PERIODS
  timeperiods_names <-
    txt[(xtimeperiods + 1)] %>% str_split_fixed(",", 2) %>% as.vector()
  
  timeperiods_tbl <-
    tibble(
      x1 = txt[(xtimeperiods + 2):(xnotes - 1)] %>%
        str_extract("^[0-9]+,") %>% str_remove(","),
      x2 = txt[(xtimeperiods + 2):(xnotes - 1)] %>%
        str_extract(",[[:alnum:]]+$") %>% str_remove_all(","),
      
    ) %>%
    rename_with(~ timeperiods_names)
  
  # DATA
  data_names <-
    txt[(xdata + 1)] %>% str_split_fixed(",", 7) %>% as.vector()
  
  data_tbl <-
    as_tibble(txt[(xdata + 2):(length(txt))] %>%
                str_split_fixed(",", 7),
              .name_repair = ~ data_names)
  
  # UPDATED
  updated_time <- txt[(xupdated + 1)] %>%
    str_remove('^\"') %>%
    str_remove('\"$') %>%
    strptime("%A, %d-%b-%y %H:%M:%S EDT")
  
  return(list(categories = categories_tbl, timeperiods = timeperiods_tbl, data = data_tbl, update = updated_time))
}

################################
# Check to see if data matches #
################################
check_data <- function(catcode, catlabel, scode, pcode, xmarts=marts, xmrts=mrts, xretail=retail_data, xprice = price_data) {
  
  adv_sales <- xmarts$data %>%
    filter(
      cat_idx == xmarts$categories %>%  filter(cat_code == catcode) %>% pull(cat_idx),
      dt_idx == "1",
      # sales monthly $mln
      et_idx == "0",
      geo_idx == "1",
      is_adj == "1"
    ) %>%
    left_join(xmarts$timeperiods, by = 'per_idx') %>%
    transmute(
      month = yearmonth(per_name, "%b%Y"),
      value = parse_number(val),
      key = "advance"
    ) %>%
    as_tsibble(index = month,
               regular = TRUE,
               key = key)
  
  sales <- xmrts$data %>%
    filter(
      cat_idx == xmrts$categories %>%  filter(cat_code == catcode) %>% pull(cat_idx),
      dt_idx == "1",
      # sales monthly $mln
      et_idx == "0",
      geo_idx == "1",
      is_adj == "1"
    ) %>%
    left_join(xmrts$timeperiods, by = 'per_idx') %>%
    transmute(
      month = yearmonth(per_name, "%b%Y"),
      value = parse_number(val),
      key = "actual"
    ) %>%
    as_tsibble(index = month,
               regular = TRUE,
               key = key)
  
  bea_sales <- xretail %>%
    filter(SeriesCode == scode) %>%
    transmute(
      month = yearmonth(TimePeriod), value = DataValue, key = 'bea') %>% 
    as_tsibble(index = month, regular = TRUE, key = key)
  
  prices <- xprice %>%
    filter(SeriesCode == pcode) %>%
    transmute(
      month = yearmonth(TimePeriod),
      value = (DataValue/lag(DataValue)-1)*100,
      key = "price_mom") %>%
    as_tsibble(index = month, regular = TRUE, key = key)
  
  
  # inventory
  inventories <- mrts$data %>%
    filter(
      cat_idx == mrts$categories %>%  filter(cat_code == catcode) %>% pull(cat_idx),
      dt_idx == "2", # inventory 
      et_idx == "0",
      geo_idx == "1",
      is_adj == "1"
    ) %>%
    left_join(mrts$timeperiods, by = 'per_idx') %>%
    transmute(
      month = yearmonth(per_name, "%b%Y"),
      value = parse_number(val),
      key = "inv_mrts") %>%
    as_tsibble(index = month, regular = TRUE, key = key)
  
  
  bind_rows(adv_sales, sales, bea_sales, prices, inventories) %>%
    filter_index("2021 Jul" ~ .) %>% pivot_wider(names_from = key)
}

##################################
# plot charts (with inventories) #
##################################
plot_charts <- function(catcode, catlabel, pcode, plot_is = TRUE, xmarts = marts, xmrts=mrts, xprice = price_data) {
  
  # SALES
  adv_sales <- xmarts$data %>%
    filter(
      cat_idx == xmarts$categories %>%  filter(cat_code == catcode) %>% pull(cat_idx),
      dt_idx == "1",
      # sales monthly $mln
      et_idx == "0",
      geo_idx == "1",
      is_adj == "1"
    ) %>%
    left_join(xmarts$timeperiods, by = 'per_idx') %>%
    transmute(
      month = yearmonth(per_name, "%b%Y"),
      value = parse_number(val),
      key = "advance"
    ) %>%
    as_tsibble(index = month,
               regular = TRUE,
               key = key)
  
  i_s <- xmrts$data %>%
    filter(
      cat_idx == xmrts$categories %>%  filter(cat_code == catcode) %>% pull(cat_idx),
      dt_idx == "3",
      # inventory / sales ratio
      et_idx == "0",
      geo_idx == "1",
      is_adj == "1"
    ) %>%
    left_join(xmrts$timeperiods, by = 'per_idx') %>%
    transmute(month = yearmonth(per_name, "%b%Y"), value = parse_number(val)) %>%
    as_tsibble(index = month, regular = TRUE)
  
  # PRICES
  prices <- xprice %>%
    filter(SeriesCode == pcode) %>%
    transmute(month = yearmonth(TimePeriod), value = DataValue) %>%
    as_tsibble(index = month, regular = TRUE)
  
  # CHART
  p1 <-
    left_join(adv_sales,
              prices,
              by = 'month',
              suffix = c(".sales", ".price")) %>%
    mutate(volume.sales = value.sales / value.price * value.price[n()]) %>%
    select(-value.price,-key) %>%
    pivot_longer(-month) %>%
    filter_index("2018 Jan" ~ .) %>%
    ggplot(aes(month, value)) +
    scale_x_yearmonth(labels = scales::label_date_short(),
                      breaks = scales::breaks_pretty(n = 10)) +
    scale_y_continuous(
      breaks = scales::breaks_pretty(n = 10),
      labels = scales::label_dollar(scale = 1e-3, accuracy = 1),
      trans = 'log',
      sec.axis = dup_axis()
    ) +
    geom_line(size = 1) +
    labs(title = "Sales (USD bn, monthly)",
         x = NULL,
         y = NULL) +
    facet_wrap(vars(name), scales = "fixed")
  
  
  p2 <- prices %>%
    mutate(value = value / lag(value, 12) - 1) %>%
    ggplot(aes(month, value)) +
    scale_x_yearmonth(labels = scales::label_date_short(),
                      breaks = scales::breaks_pretty(n = 15)) +
    scale_y_continuous(breaks = scales::breaks_pretty(n = 10),
                       labels = scales::label_percent()) +
    geom_line(size = 1) +
    labs(title = "Prices YoY",
         x = NULL,
         y = NULL)
  
  p3 <- i_s %>%
    ggplot(aes(month, value)) +
    scale_x_yearmonth(labels = scales::label_date_short(),
                      breaks = scales::breaks_pretty(n = 15)) +
    scale_y_continuous(breaks = scales::breaks_pretty(n = 10)) +
    geom_line(size = 1) +
    labs(title = "Inventories / Sales Ratio (in months)",
         x = NULL,
         y = NULL)
  
  if (plot_is) {p1 / (p2 + p3) + plot_annotation(title = catlabel)}
  else {p1 / p2 + plot_annotation(title = catlabel)}
}


########################################
# Add forecast prices for retail sales #
########################################
add_new_price <- function(scode, cpi_goods, p_data) {
  
  db <- cpi_goods %>%
    pivot_wider(names_from = series_id, values_from = value) %>%
    full_join(
      p_data %>%
        filter(SeriesCode == scode) %>%
        transmute(month = yearmonth(TimePeriod), value = DataValue),
      by = 'month'
    ) %>%
    mutate(across(-month, ~ .x / lag(.x) - 1)) %>%
    filter_index("1992 Feb" ~ .) %>%
    as_tibble()

  # Elimina as obs NAs para o 'preco'
  idx <- !is.na(db$value)
  
  # Estimate model with all ncomps
  model <-
    plsr(
      value ~ .,
      ncomp = 50,
      data = select(db[idx, ],-month),
      scale = TRUE,
      validation = "LOO"
    )
  
  #validationplot(model)
  # Estimate number of components
  ncomp.onesigma <-
    selectNcomp(model, method = "onesigma", plot = FALSE)
  
  cat("\n n. de componentes: ", ncomp.onesigma, "\n")
  # Estimate model with recommended number of components
  model <-
    plsr(
      value ~ .,
      ncomp = ncomp.onesigma,
      data = select(db[idx, ],-month),
      scale = TRUE,
      validation = "LOO"
    )
  
  # check fit
  # plot(
  #   model,
  #   ncomp = 1:ncomp.onesigma,
  #   asp = 1,
  #   line = TRUE
  # )
  
  last_price <-
    filter(p_data,
           SeriesCode == scode,
           TimePeriod == zoo::as.yearmon("Mar 2022")) %>%
    pull(DataValue)
  
  price_fcast <-
    last_price * (1 + drop(predict(model, newdata = db[!idx, ]))[ncomp.onesigma])
  
  cat("\n Inflation: ", (price_fcast / last_price - 1)*100, "\n")
  
  p_data <- p_data %>%
    add_case(
      SeriesCode = scode,
      TimePeriod = zoo::as.yearmon("Apr 2022"),
      DataValue = price_fcast
    )
  return(p_data)
}

```

```{r download_data}

# Download retail sales data 
DOWNLOAD = FALSE

if (DOWNLOAD) {
  download.file(
    "https://www.census.gov/econ/currentdata/datasets/MARTS-mf.zip",
    "data/us-retail-sales/MARTS-mf.zip",
    mode = "wb"
  )
  download.file(
    "https://www.census.gov/econ/currentdata/datasets/MRTS-mf.zip",
    "data/us-retail-sales/MRTS-mf.zip",
    mode = "wb"
  )
}

unzip("data/us-retail-sales/MARTS-mf.zip", exdir = "data/us-retail-sales")
unzip("data/us-retail-sales/MRTS-mf.zip", exdir = "data/us-retail-sales")

# extract tables from files

marts <- read_lines("data/us-retail-sales/MARTS-mf.csv", skip_empty_rows = TRUE, lazy = TRUE)
mrts  <- read_lines("data/us-retail-sales/MRTS-mf.csv", skip_empty_rows = TRUE, lazy = TRUE)

marts <- extract_tables(marts)
mrts <- extract_tables(mrts)

# Download PRICES and Retail data (from BEA)
myboard <- pins::board_folder("D:\\RData\\BEA")
price_data <- pins::pin_read(myboard, "U00700_M")
retail_data <- pins::pin_read(myboard, "U00500_M")
inventory_data <- pins::pin_read(myboard, "U001BC_M")

# this dataframe will receive price forecasts
new_price_data <- price_data

# Get CPI prices
myboard <- pins::board_folder("D:\\RData")
cpi <- pins::pin_read(myboard, "USCPI")
cpi_item <- pins::pin_read(myboard, "USCPI-ITEM")
cpi_series <- pins::pin_read(myboard, "USCPI-SERIES")


cpi_goods_item <- read_csv("cpi-item.csv", col_types = cols(item_code = col_character(), 
    item_name = col_character(), display_level = col_integer(), 
    sort_sequence = col_integer(), select = col_character()))

cpi_goods_ticker_sa <- 
  cpi_series %>% 
  filter(
    area_code == "0000",
    item_code %in% (cpi_goods_item %>% filter(select == "x") %>% pull(item_code)),
    seasonal == "S",
    periodicity_code == "R"
  ) %>% pull(series_id)

cpi_goods_ticker_nsa <- 
  cpi_series %>% 
  filter(
    area_code == "0000",
    item_code %in% (cpi_goods_item %>% filter(select == "x") %>% pull(item_code)),
    seasonal == "U",
    periodicity_code == "R"
  ) %>% pull(series_id)

cpi_goods_ticker_nsa <- base::setdiff(cpi_goods_ticker_nsa, str_replace_all(cpi_goods_ticker_sa, "CUSR", "CUUR"))

# CPI goods contain all the CPI series used to construct PCE / retail sales deflators
cpi_goods <- cpi %>% 
  filter(series_id %in% c(cpi_goods_ticker_sa, cpi_goods_ticker_nsa)) %>% 
  select(month, series_id, value) %>% 
  filter_index("1992 Jan" ~. )

```

# Main aggregates

## Introduction

## Retail and food services (total)
```{r}
check_data("44X72", "RETAIL & FOOD SERVICES", "T05001", "T07001")

new_price_data <- add_new_price("T07001", cpi_goods, new_price_data)

plot_charts("44X72", "RETAIL & FOOD SERVICES", "T07001", FALSE)
plot_charts("44X72", "RETAIL & FOOD SERVICES", "T07001", FALSE, xprice = new_price_data)

```

## Retail 
```{r}
check_data("44000", "RETAIL TOTAL (EXCL. FOOD SERVICES)", "T05002", "T07002")
plot_charts("44000", "RETAIL TOTAL (EXCL. FOOD SERVICES)", "T07002")

new_price_data <- add_new_price("T07002", cpi_goods, new_price_data)



```

## Retail total excl. motor vehicle, parts and gasoline
```{r}
check_data("44W72", "RETAIL & FOOD SERVICES(EXCL. MOTOR VEH & PARTS AND GASOLINE STATIONS)", "T05046", "T07046")


new_price_data <- add_new_price("T07046", cpi_goods, new_price_data)
plot_charts("44W72", "RETAIL & FOOD SERVICES(EXCL. MOTOR VEH & PARTS AND GASOLINE STATIONS)", "T07046", FALSE, xprice = new_price_data)

```


# Retail by kind of business

## Motor vechicle and parts dealers
```{r}
check_data("441", "MOTOR VEHICLE & PARTS DEALERS", "T05003", "T07003")
plot_charts("441", "MOTOR VEHICLE & PARTS DEALERS", "T07003")

new_price_data <- add_new_price("T07003", cpi_goods, new_price_data)
plot_charts("441", "MOTOR VEHICLE & PARTS DEALERS", "T07003", xprice = new_price_data)


```


## Furniture
```{r}
check_data("442", "FURNITURE & HOME FURNISHING STORES", "T05008", "T07008")
plot_charts("442", "FURNITURE & HOME FURNISHING STORES", "T07008", FALSE)

new_price_data <- add_new_price("T07008", cpi_goods, new_price_data)
plot_charts("442", "FURNITURE & HOME FURNISHING STORES", "T07008", FALSE, xprice = new_price_data)
```

## Electronic and appliances stores
```{r}
check_data("443", "ELECTRONICS & APPLIANCE STORES", "T05011", "T07011")

new_price_data <- add_new_price("T07011", cpi_goods, new_price_data)
plot_charts("443", "ELECTRONICS & APPLIANCE STORES", "T07011", FALSE, xprice = new_price_data)



```

## Building material and garden equipment
```{r}
check_data("444", "BUILDING MATERIALS, GARDEN EQ. & SUPPLY DEALERS", "T05012", "T07012")

new_price_data <- add_new_price("T07012", cpi_goods, new_price_data)
plot_charts("444", "BUILDING MATERIALS, GARDEN EQ. & SUPPLY DEALERS", "T07012", xprice = new_price_data)


```

## Food and beverage stores
```{r}
check_data("445", "FOOD & BEVERAGE STORES", "T05015", "T07015")

new_price_data <- add_new_price("T07015", cpi_goods, new_price_data)
plot_charts("445", "FOOD & BEVERAGE STORES", "T07015", xprice = new_price_data)


```

## Health and personal care stores
```{r}
check_data("446", "HEALTH & PERSONAL CARE STORES", "T05027", "T07027")

new_price_data <- add_new_price("T07027", cpi_goods, new_price_data)
plot_charts("446", "HEALTH & PERSONAL CARE STORES", "T07027", FALSE, xprice = new_price_data)


```

## Gasoline stations
```{r}
check_data("447", "GASOLINE STATIONS", "T07028", "T07028")

new_price_data <- add_new_price("T07028", cpi_goods, new_price_data)
plot_charts("447", "GASOLINE STATIONS", "T07028", FALSE, xprice = new_price_data)

```

## Clothing and accessories stores
```{r}
check_data("448", "CLOTHING & CLOTHING ACCESSORIES STORES", "T05019", "T07019")

new_price_data <- add_new_price("T07019", cpi_goods, new_price_data)
plot_charts("448", "CLOTHING & CLOTHING ACCESSORIES STORES", "T07019", xprice = new_price_data)


     
```

## Sporting goods, hobby, books, musical instruments
```{r}
check_data("451", "SPORTING GOODS, HOBBY, MUSICAL INST. & BOOK STORES","T05029", "T07029" )

new_price_data <- add_new_price("T07029", cpi_goods, new_price_data)
plot_charts("451", "SPORTING GOODS, HOBBY, MUSICAL INST. & BOOK STORES", "T07029", FALSE, xprice = new_price_data)


```

## General merchandise stores
```{r}
check_data("452", "GENERAL MERCHANDISE STORES", "T05023", "T07023")

new_price_data <- add_new_price("T07023", cpi_goods, new_price_data)
plot_charts("452", "GENERAL MERCHANDISE STORES", "T07023", xprice = new_price_data)


```

## Miscellaneous store retailers
```{r}
check_data("453", "MISCELLANEOUS STORES RETAILERS", "T05032", "T07032")

new_price_data <- add_new_price("T07032", cpi_goods, new_price_data)
plot_charts("453", "MISCELLANEOUS STORES RETAILERS", "T07032", FALSE, xprice = new_price_data)


```

## Nonstore retailers
```{r}
check_data("454", "NONSTORE RETAILERS", "T05037", "T07037")

new_price_data <- add_new_price("T07037", cpi_goods, new_price_data)
plot_charts("454", "NONSTORE RETAILERS", "T07037", FALSE, xprice = new_price_data)


```

## Food services and drinking places
```{r}
check_data("722", "FOOD SERVICES & DRINKING PLACES", "T05041", "T07041")

new_price_data <- add_new_price("T07041", cpi_goods, new_price_data)
plot_charts("722", "FOOD SERVICES & DRINKING PLACES", "T07041", FALSE, xprice = new_price_data)


```


